---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis-CI Build Status](https://travis-ci.org/benmarwick/smps.svg?branch=master)](https://travis-ci.org/benmarwick/smps)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/README-",
  cache = FALSE, 
  message = FALSE,
  warning = FALSE
)
library(smps)
```

# SMPS

## Introduction

This package produces time series colour contour plots of data from [Scanning Mobility Particle Sizer (SMPS)](https://en.wikipedia.org/wiki/Scanning_mobility_particle_sizer) data. The SMPS is the standard method to measure airborne particle size distributions. The kind of plots that this package produces are widely used for visualising SMPS data in atmospheric science. 

## How to install the package

You will need the `devtools` package from CRAN so you can install this package from github with the following line of code:

```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("benmarwick/smps")
library(smps)
```

## How to use the package

First, read your data into R using `read.csv` or similar, and check that it is in a format that works with this package. 

You can inspect the example dataset that comes with the package like this:

```{r eval=FALSE}
data("my_data")
str(my_data)
head(my_data)
View(my_data)
```

Second, prepare your data for plotting using the `prepare_data` function. This converts the variable types and data format into forms suitable for plotting. It also does interpolation of the measured values to give a smooth contour plot. The interpolation can take a minute or two. You can experiment with the interpolation parameters to speed up the operation.

```{r, cache = TRUE}
# prepare the data
prepared_data <- prepare_data(my_data)
```

Third, plot the data, like so:

```{r prep}
smps_plot(prepared_data) 
```

We can also have a log scale on the y-axis:

```{r log}
smps_plot(prepared_data, y_axis = "log")
```

We can plot a data from a single day:

```{r oneday}
smps_plot(prepared_data, 
          day = "2013-01-26", 
          aspect_ratio = 1/2, 
          h = '3 hours')
```

We can customise a variety of options, including the colour ramp:

```{r options}
require(viridis)
smps_plot(prepared_data, 
          font_size = 9, 
          h = '3 hours',
          y_breaks = 100,
          aspect_ratio = 1/4,
          colour_ramp = viridis) 
```

Because the output is a ggplot object, we can add typical ggplot elements to it like a regular `ggplot()` call:

```{r ggpl}
require(ggplot2)
smps_plot(prepared_data, font_size = 8) + 
  xlab("Three days of interesting data") +
  theme(legend.key.size = unit(0.1, "in"))
```

With `gridExtra` we can stack a few plots together, and arrange a common legend to the side:

```{r, combine, cache = TRUE}
library(gridExtra)
# select the days to plot
the_days <- c("2013-01-26", "2013-01-27", "2013-01-28")
# draw three plots, one day each, collect them in a list
the_plots <- lapply(the_days, function(i) smps_plot(prepared_data, 
                                                    day = i, 
                                                    y_axis = "log",  
                                                    h = '3 hours', 
                                                    font_size = 8) +   
                                                    guides(fill = FALSE))
  
# we extract the legend of the a plot to use separately
the_legend <- get_legend(smps_plot(prepared_data, 
                                   day = "2013-01-26", 
                                   font_size = 10))

# arrange the plots and scale
grid.arrange(the_plots[[1]], 
             the_plots[[2]], 
             the_plots[[3]], 
             the_legend, 
             layout_matrix = rbind(c(1,1,4),c(2,2,4),c(3,3,4)))
```


Please note that this project is released with a [Guide to Contributing](CONTRIBUTING.md) and a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

